---
title: "PCWG_share_01_main"
author: "Andy Clifton"
output: html_document
---
<!-- ## Reading the source code?
This is R Markdown. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

This script is designed to be used with RStudio. When you click the **Knit** button in RStudio a document will be generated that includes the output of any embedded R code chunks within the document. -->



# Introduction
This document contains the results of the Power Curve Working Group's Share_01 exercise, which ran from October to December 2015. The document and results are generated using the programing language `R` from the _PCWG_share_01_main.rmd_ file and can be run by participants themselves.

## How to use PCWG_share_01_main.rmd
install R (<http://www.r-project.org>) and Rstudio (<http://www.rstudio.com>), and then create a directory with all of the code and files (see below). When you click the **Knit** button in RStudio a document will be generated that includes text and results from the code embedded in _PCWG_share_01_main.rmd_.

```{r, echo = FALSE}
rm(list = ls())
```
## User Inputs
The _project.root_ variable defines the location of the files required for this analysis. The _made.by_ variable forms part of a label that will be added to the plots. _data.public_ is a flag that indicates whether the results of the analysis are intended to be public, or not. 

```{r}
# Where can files be found?
project.root <- file.path('/Users/aclifton/Dropbox/work/A2e PRUF Working Folder',
                          'FY15 Activities/PCWG Share 01')

# Who ran this script
made.by = "A. Clifton, NREL"

# Will data be public or not?
data.public = TRUE

# Reanalyze existing data?
data.reanalyze = FALSE
```

## Directory structure
The folowing files should be placed in the _project.root_ directory:

* PCWG_share_01_main.Rmd
* /__analysis__ directory containg results of the analysis
* /__code__ directory containing functions required for the analysis
* /__data__ directory containing all data files to be analyzed. This can include further sub directories. All .xls files contained in __data__ and sub directories will be used in the analysis.

## Packages
This script requires the _ggplot2_, _grid_, _knitr_, and _XLConnect_ packages to run. These are called from the script but you may need to install them directly. For details of how to install packages, see the RStudio help. 
```{r, message=FALSE, echo = FALSE}
require(ggplot2)
require(ggthemes)
require(grid)
require(XLConnect)
require(knitr)
require(reshape2)

# define the working directory
working.dir <- project.root
setwd(working.dir)

#identify data directory
data.dir = file.path(project.root,
                     "data")

# define where functions live
code.dir = file.path(project.root,
                     "code")

# source these functions
code.files = dir(code.dir, pattern = "\\.R$")
for (file in code.files){
  source(file = file.path(code.dir,file))
}

# configure graphics appearance
theme_set(theme_PCWG(base_size = 8, 
                     base_family = "sans"))

```

# Results from each data set
We now analyse the data from each data set. The plots are saved to their own directories in the _analysis_ directory. If _data.public_ is FALSE, plots will be created for every data file. If _data.public_ is TRUE, only the final, aggregated data plots will be created.
```{r, results='asis', echo = FALSE}

if (data.reanalyze == TRUE){
  # identify the data sets that we have available
  data.files = dir(data.dir, 
                   recursive = TRUE,
                   pattern = "[.xls]")
  
  # create an empty list to store data in at a later date 
  all.data <- list(sub = NULL,
                   meta = NULL,
                   errors = NULL)
  for (data.file in data.files){
    # Read this data set
    wb <- loadWorkbook(file.path(data.dir,data.file))
    in.sub <- ReadSubmissionData(wb)
    in.sub$data.file <- data.file
    in.meta <- ReadMetaData(wb)
    in.meta$data.file <- data.file
    # get *all* errors
    in.errors <- ReadErrorData(wb)
    
    if (data.public == FALSE){
      # write out results from this case to file and to the document  
      cat(text = "\n## Data set", in.sub$random.ID, "from", data.file, "\n")
      # check to see if we have a directory for this case
      output.dir = file.path(project.root, 
                             "analysis",
                             strsplit(data.file, "\\.")[[1]][1])
      dir.create(output.dir, showWarnings = FALSE,recursive = TRUE)
      
      # create a label we'll use to annotate plots
      plot.label <- labelSubmission(in.sub,
                                    made.by)
      # plot errors
      PlotSubErrorsByWS(in.errors$by.WS,
                        plot.label,
                        output.dir)
      PlotSubErrorsByTOD(in.errors$by.TOD,
                         plot.label,
                         output.dir)
      PlotSubErrorsByCM(in.errors$by.CM,
                        plot.label,
                        output.dir)
      PlotSubErrorsByWD(in.errors$by.WD,
                        plot.label,
                        output.dir)
      PlotSubErrorsBy4CM(in.errors$by.4CM,
                         plot.label,
                         output.dir)
      
      # dummy text to clear knitr buffers  
      cat(text = "\n")
    }
    
    # prepare data for aggregation
    all.data <- aggregateDataSets(all.data,
                                  in.sub,
                                  in.meta,
                                  in.errors)
  }
}
```

## Turbine Sizes, Characteristics, and Locations

```{r, echo=FALSE}
# define the path to the directory that we will use
output.dir = file.path(project.root, 
                       "analysis",
                       "all")
dir.create(output.dir, 
           showWarnings = FALSE,
           recursive = TRUE)

# save the data
if (data.reanalyze == TRUE){
  save(list = c("project.root",
                "made.by",
                "data.public",
                "working.dir",
                "output.dir",
                "all.data"), 
       file = file.path(output.dir,"AggregatedData.RData"),
       envir = .GlobalEnv)
} else {
  load(file = file.path(output.dir,"AggregatedData.RData"))
}

# create a label we'll use for the plots
plot.label <- labelAggregate(as.character(NROW(all.data$sub)),
                             made.by)
```

In total, `r nrow(all.data$sub)` data sets were submitted.

The `r nrow(all.data$sub)` data sets include tests carried out in the period from `r min(all.data$meta$year.of.measurement)` to `r max(all.data$meta$year.of.measurement)`. Turbine diameters range from `r min(all.data$meta$turbine.dia)` to `r max(all.data$meta$turbine.dia)` m, while hub heights range from `r min(all.data$meta$turbine.height)` to `r max(all.data$meta$turbine.height)` m.

```{r, echo=FALSE}
PlotAllTurbineYearSizeDia(all.data$meta,
                          plot.label,
                          output.dir)
```

## Test Characteristics
```{r, echo=FALSE}
PlotAllTurbineLocations(all.data$meta,
                        plot.label,
                        output.dir)

MapAllTurbineLocations(all.data$meta,
                       plot.label,
                       code.dir,
                       output.dir)

```

# Aggregated results
In this section, data from all of the individual submissions have been combined. 

## Baseline Inner and Outer Range Error Histograms

```{r, echo = FALSE}
PlotAllBaselineErrorsByRange(all.data$errors$by.Range,
                             plot.label,
                             output.dir)
```

## Comparison of inner and outer range errors

plot baseline inner range versus other inner range NME, color by method


## The Effect of The Turbulence Correction on the Outer Range Error
```{r, echo = FALSE}
PlotAllOuterErrorsTurbulenceCorrection(all.data$errors$by.Range,
                                       plot.label,
                                       output.dir)
```

## Errors versus wind speed for all submissions

```{r, echo=FALSE}
PlotAllErrorsByWSBin_Lines(all.data$errors$by.WS,
                           plot.label,
                           output.dir)
```               

## Errors Binned by Wind Speed

```{r, echo=FALSE}
PlotAllErrorsByWSBin_BoxNWhiskers(all.data$errors$by.WS,
                                  plot.label,
                                  output.dir)
```               

## Change in errors binned by Wind Speed
```{r, echo=FALSE}
# work out the change compared to the baseline in each data set
temp <- NULL
for (file in unique(all.data$errors$by.WS$data.file)){
  sub <- all.data$errors$by.WS[all.data$errors$by.WS$data.file == file,]
  for (ds in unique(sub$correction)){
    for (WS.cell in unique(sub$x.min)){
      # get the reference NME and NMAE
      baseline <- sub[(sub$correction == "Baseline") & 
                        (sub$x.min == WS.cell),]
      # get the NME and NMAE for this cell
      new <- sub[(sub$correction == ds) & 
                   (sub$x.min == WS.cell),]
      delta <- data.frame(new,
                          error.delta.pc = c(new$error.val.pc[new$error.name == "NME"] - baseline$error.val.pc[baseline$error.name == "NME"],
                                             new$error.val.pc[new$error.name == "NMAE"] - baseline$error.val.pc[baseline$error.name == "NMAE"]))
      temp <- rbind(temp,
                    delta)
    }
  }
}

PlotAllChangeInErrorsByWSBin(temp,
                             plot.label,
                             output.dir)

remove(temp)
```

## Errors Binned by Wind Speed and Ti
The plot



```{r, echo=FALSE}
# work out the change compared to the baseline in each data set
temp <- NULL
for (file in unique(all.data$errors$by.4CM$data.file)){
  sub <- all.data$errors$by.4CM[all.data$errors$by.4CM$data.file == file,]
  for (ds in unique(sub$correction)){
    for (WS.cell in unique(sub$WS.cell)){
      for (Ti.cell in unique(sub$Ti.cell)){
        # get the reference NME and NMAE
        baseline <- sub[(sub$correction == "Baseline") & 
                          (sub$WS.cell == WS.cell) &
                          (sub$Ti.cell == Ti.cell),]
        # get the NME and NMAE for this cell
        new <- sub[(sub$correction == ds) & 
                     (sub$WS.cell == WS.cell) &
                     (sub$Ti.cell == Ti.cell),]
        delta <- data.frame(new,
                            error.delta.pc = c(new$error.val.pc[new$error.name == "NME"] - baseline$error.val.pc[baseline$error.name == "NME"],
                                               new$error.val.pc[new$error.name == "NMAE"] - baseline$error.val.pc[baseline$error.name == "NMAE"]))
        temp <- rbind(temp,
                      delta)
      }
    }
  }
}

PlotAllChangeInErrorsBy4CM(temp,
                           plot.label,
                           output.dir)
```               
